\cchapter{Synchronization}{synchronization}
\label{chap:synchronization}

The \kcode{barrier} construct is a stand-alone directive that requires all threads
of a team (within a contention group) to execute the barrier and complete
execution of all tasks within the region, before continuing past the barrier.

The \kcode{critical} construct is a directive that contains a structured block. 
The construct allows only a single thread at a time to execute the region.
Multiple \kcode{critical} regions may exist in a parallel region, and they may act
cooperatively (only one thread at a time in all \kcode{critical} regions) or
separately (only one thread at a time in each \kcode{critical} regions when a
unique name is supplied on each \kcode{critical} construct). An optional (lock)
\kcode{hint} clause may be specified on a named \kcode{critical} construct to
provide the OpenMP runtime guidance in selecting a locking mechanism.

On a finer-grained scale, the \kcode{atomic} construct allows only a single
thread at a time to have atomic access to a storage location involving a single
atomic read, write, or update statement, and a limited number of combinations
when specifying the \kcode{capture} and \kcode{compare} \plc{extended-atomic}
clauses.  Unlike the \kcode{read} and \kcode{write} \plc{atomic} clauses, the
\kcode{update} \plc{atomic} clause is optional and implied for update
statements if not explicitly specified.. The \plc{memory-order} clause can be
used to specify the degree of memory ordering enforced by an \kcode{atomic}
construct. From weakest to strongest, they are \emph{relaxed} (the default),
\emph{acquire} and/or \emph{release} (specified with \kcode{acquire},
\kcode{release}, or \kcode{acq_rel} clauses), and \emph{sequentially consistent}
(specified with \kcode{seq_cst}).  Please see the details in the \docref{atomic
Construct} subsection of the \docref{Synchronization Constructs and Clauses}
chapter in the OpenMP Specification document.

The \kcode{ordered} construct may be block-associated or stand-alone. The
block-associated form may appear in worksharing-loop (\kcode{for} or
\kcode{do}), \kcode{simd}, or worksharing-loop SIMD (\kcode{for simd} or
\kcode{do simd}) region. It sequentializes and orders the execution of the
\kcode{ordered} regions while allowing code outside the region to run in
parallel. The stand-alone \kcode{ordered} construct specifies cross-iteration
dependences in a \emph{doacross} loop nest. The \kcode{doacross} clause uses a
\kcode{sink} \plc{dependence-type}, along with an iteration vector argument
(\plc{vec}) to indicate the iteration that satisfies the dependence.  The
\kcode{doacross} clause with a \kcode{source} \plc{dependence-type} (and
optional \kcode{omp_curr_iteration} keyword as the iteration vector
representing the current iteration) specifies dependence satisfaction.

The \kcode{flush} directive is a stand-alone construct for enforcing consistency
between a thread's view of memory and the view of memory for other threads (see
the Memory Model chapter of this document for more details). When the construct
is used with an explicit variable list, a \emph{strong flush} that forces a
thread's temporary view of memory to be consistent with the actual memory is
applied to all listed variables. When the construct is used without an explicit
variable list and either without a \plc{memory-order} clause or with the
\kcode{seq_cst} \plc{memory-order} clause, a strong flush is applied to all
locally thread-visible data as defined by the base language, and additionally
the construct provides both acquire and release memory ordering semantics.
When an explicit variable list is not present and a \plc{memory-order} clause
other than \kcode{seq_cst} is present, the construct provides acquire and/or
release memory ordering semantics according to the \plc{memory-order} clause,
but no strong flush is performed. A resulting strong flush that applies to a
set of variables effectively ensures that no memory (load or store) operation
for the affected variables may be reordered across the \kcode{flush} directive.

General-purpose routines provide mutual exclusion semantics through locks, 
represented by lock variables.  
The semantics allows a task to \emph{set}, and hence 
\emph{own} a lock, until it is \emph{unset} by the task that set it. A 
\emph{nestable} lock can be set multiple times by a task, and is used
when in code requires nested control of locks.  A \emph{simple lock} can
only be set once by the owning task. There are specific calls for the two
types of locks, and the variable of a specific lock type cannot be used by the
other lock type.  

Any explicit task will observe the synchronization prescribed in a
\kcode{barrier} construct and an implied barrier.  Also, additional
synchronizations are available for tasks.  A task will wait at a
\kcode{taskwait} for all of its child tasks to complete.  A \kcode{taskgroup}
construct creates a region in which the current task is suspended at the end of
the region until all child tasks, and their descendants, have completed.
Scheduling constraints on task execution can be prescribed by the
\kcode{depend} clause to enforce a dependence on previously generated tasks.
More details on controlling task executions can be found in the \docref{Tasking
Constructs} chapter in the OpenMP Specifications document. %(DO REF. RIGHT.)


%===== Examples Sections =====
\input{synchronization/critical}
\input{synchronization/worksharing_critical}
\input{synchronization/barrier_regions}
\input{synchronization/atomic}
\input{synchronization/atomic_cas}
\input{synchronization/atomic_restrict}
\input{synchronization/atomic_hint}
\input{synchronization/acquire_release}
\input{synchronization/ordered}
\input{synchronization/depobj}
\input{synchronization/doacross}
\input{synchronization/locks}
\input{synchronization/init_lock}
\input{synchronization/init_lock_with_hint} 
\input{synchronization/lock_owner}
\input{synchronization/simple_lock}
\input{synchronization/nestable_lock}
\input{synchronization/safesync}

