%\pagebreak
\section{Structure Mapping}
\label{sec:structure_mapping}
\index{mapping!structure}

\index{directives!begin declare target@\kcode{begin declare target}}
\index{begin declare target directive@\kcode{begin declare target} directive}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In the example below, only structure elements \ucode{S.a}, \ucode{S.b} and
\ucode{S.p} of the \ucode{S} structure appear in \kcode{map} clauses of a
\kcode{target} construct. Only these components have corresponding variables
and storage on the device.  Hence, the large arrays, \ucode{S.buffera} and
\ucode{S.bufferb}, and the \ucode{S.x} component have no storage on the device
and cannot be accessed.

Also, since the pointer member \ucode{S.p} is used in an array section
(\ucode{S.p[:N]}) of a \kcode{map} clause, the pointer member
\ucode{S.p} on the device is \emph{attached} to the array storage of
the array section on the device. Explicitly mapping the pointer member
\ucode{S.p} is optional in this case.

Note: The buffer arrays and the \ucode{x} variable have been grouped together,
so that the components that will reside on the device are all together (without
gaps). This allows the runtime to optimize the transfer and the storage
footprint on the device.

\cexample[5.1]{target_struct_map}{1}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The following example is a slight modification of the above example for a C++
class.  In the member function \ucode{SAXPY::driver} the pointer member
\ucode{p} on the device is attached to the array section \ucode{p[:N]} on the
device.

\cppexample[5.1]{target_struct_map}{2}

The next example shows two ways in which the structure may be
\emph{incorrectly} mapped.

In Case 1, the array section \ucode{S1.p[:N]} is first mapped in an enclosing
\kcode{target_data} construct, and the \kcode{target} construct then implicitly
maps the structure \ucode{S1}. The initial map of the array section does not
map the base pointer \ucode{S1.p} -- it only maps the elements of the array
section.  Furthermore, the implicit map is not sufficient to ensure pointer
attachment for the structure member \ucode{S1.p} (refer to the conditions for
pointer attachment described in Section~\ref{sec:pointer_mapping}).
Consequentially, the dereference of \ucode{S->p} in the call to
\ucode{saxpyfun} will probably fail because \ucode{S->p} contains a host
address.

In Case 2, again an array section is mapped on an enclosing \kcode{target data}
construct. This time, the nested \kcode{target} construct explicitly maps
\ucode{S2.p}, \ucode{S2.a}, and \ucode{S2.b}. But as in Case 1, this does not
satisfy the conditions for pointer attachment since the construct must map a
list item for which \ucode{S2.p} is a base pointer, and it must do so when
\ucode{S2.p} is already present on the device or will be created on the device
as a result of the same construct.


\cexample[5.1]{target_struct_map}{3}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The following example correctly implements pointer attachment cases that
involve implicit structure maps.

In Case 1, members \ucode{p}, \ucode{a}, and \ucode{b} of the structure \ucode{S1}
are explicitly mapped by the \kcode{target_data} construct, to avoid
mapping parts of \ucode{S1} that aren't required on the device. The mapped
\ucode{S1.p} is attached to the array section \ucode{S1.p[:N]}, and remains
attached while it exists on the device (for the duration of
\kcode{target_data} region).  Due to the \ucode{S1} reference inside the
nested \kcode{target} construct, the construct implicitly maps \ucode{S1} so that
the reference refers to the corresponding storage created by the enclosing
\kcode{target_data} region. Note that only the members \ucode{a},
\ucode{b}, and \ucode{p} may be accessed from this storage.

In Case 2, only the storage for the array section \ucode{S2.p[:N]} is mapped
by the \kcode{target_data} construct.  The nested \kcode{target}
construct explicitly maps \ucode{S2.a} and \ucode{S2.b} and explicitly
maps an array section for which \ucode{S2.p} is a base pointer. This satisfies
the conditions for \ucode{S2.p} becoming an attached pointer. The array
section in this case is zero-length, but the effect would be the same if the
length was a positive integer less than or equal to \ucode{N}. There is also an
implicit map of the containing structure \ucode{S2}, again due to the reference
to \ucode{S2} inside the construct. The effect of this implicit map permits
access only to members \ucode{a}, \ucode{b}, and \ucode{p}, as for Case 1.

In Case 3, there is no \kcode{target_data} construct. The \kcode{target}
construct explicitly maps \ucode{S3.a} and \ucode{S3.b} and explicitly
maps an array section for which \ucode{S3.p} is a base pointer. Again, there is
an implicit map of the structure referenced in the construct, \ucode{S3}. This
implicit map also causes \ucode{S3.p} to be implicitly mapped, because no other
part of \ucode{S3} is present prior to the construct being encountered. The
result is an attached pointer \ucode{S3.p} on the device. As for Cases 1 and 2,
this implicit map only ensures that storage for the members \ucode{a}, \ucode{b},
and \ucode{p} are accessible within the corresponding \ucode{S3} that is created
on the device.

\cexample[5.1]{target_struct_map}{4}

